### 2020-09-11 code/figure-compression.Rmd

---
title: "Figure Compression"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

This document contains code for compressing the static figure in the manuscript.

```{r}
# Load packages
library(zip)

# Save the bullet image used in the paper as a zip file
zip(
  zipfile = "../figure-static/figure-08-1.png.zip",
  files =  "../figure-static/figure-08-1.png"
)
```