### 2020-09-11 code/submission-steps.Rmd

---
title: "Sumbission Steps"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

This document contains the steps to follow to make sure all guidelines are met for submission to the ASA Data Science Journal.

1. Knit the file `paper.Rnw` (to make sure everything is up to date).

2. Run the following code to move all files necessary to the submission folder.:

    ```{r}
    # Load package
    library(zip)

    # Create a submission folder file path:
    submission_folder <- "../submission"

    # If the submission folder does not exist, create one
    if (!dir.exists(submission_folder)) dir.create(submission_folder)

    # DATA AND CODE

    # Create a temporary directory for storing code and data files for submission
    code_data_folder = "code-and-data"
    dir.create(code_data_folder)

    # Extract the R code from the files paper.Rnw to be used for submission
    knitr::purl("../paper.Rnw", output = paste0(code_data_folder, "/paper.R"))

    # Create a vector with code and data files to be included for submission 
    data_files <-
      c(
        "../readme.txt",
        "../data/bullet-test.csv",
        "../data/bullet-train.csv",
        "../data/example-signatures.csv"
      )

    # Copy the submission files to the temporary folder
    file.copy(data_files, code_data_folder)

    # Compress the folder with data and code for submission and store it in the 
    # submission folder
    zip(paste0(submission_folder, "/code-and-data.zip"), code_data_folder)

    # Remove the temporary folder with code and data for submission
    unlink(code_data_folder, recursive = TRUE)

    # FIGURES

    # Identify the folders with files for submission
    folder_eps_figures <- "../figure"
    folder_static_figures <- "../figure-static"

    # Find the EPS figure files for submission
    list_of_generated_figures <- list.files(folder_eps_figures, ".eps$", full.names = T)
    list_of_static_figures <- list.files(folder_static_figures, ".eps$", full.names = T)

    # Copy files to the submission folder
    file.copy(list_of_generated_figures, submission_folder)
    file.copy(list_of_static_figures, submission_folder)

    # LATEX FILES

    # Copy files to the submission folder
    file.copy("../paper.tex", submission_folder)
    file.copy("../paper.pdf", submission_folder)
    file.copy("../references.bib", submission_folder)
    file.copy("../WileyNJD-AMS.bst", submission_folder)
    file.copy("../WileyNJD-v2.cls", submission_folder)
    ```

3. Manually delete `figure/` and `figure-static/` from the figure file paths in `paper.tex`. For example, change `\includegraphics[width=6.5in]{figure/figure-01-1}` to `\includegraphics[width=6.5in]{figure-01-1}`.

4. Compile the file `paper.tex` (using LaTex, then bibTex, then LaTeX twice more) to make sure it compiles correctly (especially references).

5. Besides for `paper.pdf`, delete all other files generated when `paper.tex` was compiled. This includes other paper files and pdf versions of the figures.

6. Upload all remaining files to ASA Data Science Journal submission page which should be:

  - `code-and-data.zip`
  - `figure-01-1.eps` through `figure-C3-1.eps`
  - `paper.pdf`
  - `paper.tex`
  - `references.bib`
  - `WileyNJD-AMS.bst`
  - `WileyNJD-v2.cls`