### 2020-09-16 code/data-preparation.Rmd
updated data files in preparation for training the random forest model

---
title: "Bullet Data Preparation for LIME Diagnostics Paper"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center")
```

This document contains code for preparing the raw bullet data sets (training, testing, and example matching signatures) to be used the paper "Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations".

**NOTE** The code chunks where the prepared data versions are shared have `eval = FALSE`. These code chunks must be run manually. This is done to help avoid any accidental changes of the data.

Load R packages:

```{r}
library(cowplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(randomForest)
library(stringr)
library(tidyr)
library(zip)
```

Obtain features used when fitting the rtrees random forest:

```{r}
rf_features <- rownames(bulletxtrctr::rtrees$importance)
```

# Bullet Training Data

Load the raw version of the Hamby 172 and 252 bullet comparison data:

```{r}
bullet_train_raw = read.csv("../data/raw/hamby-comparisons.csv")
```

Determine the number of total study-barrel-bullet-lands in the data:

```{r}
length(unique(bullet_train_raw %>% pull(land_id1)))
```

Determine the number of total study-barrel-bullet-lands in the data when observations with tank rash are excluded:

```{r}
length(unique(bullet_train_raw %>% filter(flag == FALSE) %>% pull(land_id1)))
```

There is only one less land id when the cases with tank rash are excluded. The code below shows that this case is `Hamby252-Br9-B2-L4` since there are no rows in the filtered data without `Hamby252-Br9-B2-L4` as one of the land ids:

```{r}
length(unique(
  bullet_train_raw %>% filter(flag == TRUE) %>% filter(
    land_id1 != "Hamby252-Br9-B2-L4",
    land_id2 != "Hamby252-Br9-B2-L4"
  ) %>% pull(land_id1)
))
```

## Compare Dimensions to rtrees

The dimension of the data once the observations with tank rash (`flag != FALSE`) are removed:

```{r}
bullet_train_raw %>% filter(flag == FALSE) %>% dim()
```

The number of observations in the random forest `rtrees` (note that they do not match the number of rows in the above dataset):

```{r}
length(bulletxtrctr::rtrees$predicted)
```

The discrepancy in the number of observations has led us to decide to train a new random forest model using the data provided by Heike where `flag == FALSE` (observations known to not have tank rash). The code for the random forest can be found in the document `paper.Rnw`.

## Preparing Data

Remove any lands with tank rash, select only variables of use for the paper, add a case variable, and rename `same_source` as `samesource`:

```{r}
bullet_train <- 
  bullet_train_raw %>%
  filter(flag == FALSE) %>%
  select(-lag, -abs_lag, -overlap, -signature_length, -flag) %>%
  mutate(case = factor(1:n())) %>%
  rename("samesource" = "same_source") %>%
  select(case, land_id1, land_id2, everything())
```

Save the prepared data as a csv file:

```{r eval = FALSE}
write.csv(
  x = bullet_train, 
  file = "../data/bullet-train.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/raw/hamby-comparisons.csv.zip", 
  files = "../data/raw/hamby-comparisons.csv"
)
zip(
  zipfile = "../data/bullet-train.csv.zip", 
  files = "../data/bullet-train.csv"
)
```

# Bullet Testing Data

Load in the raw Hamby 224 datasets (sets 1 and 11):

```{r}
hamby224_set1 <- readRDS("../data/raw/h224-set1-features.rds")
hamby224_set11 <- readRDS("../data/raw/h224-set11-features.rds")
```

The bullet and land names from each of the sets are included below. This indicates that each set contains `r length(unique(hamby224_set1$bulletA))` bullets with `r length(unique(hamby224_set1$landA))` lands each.

| Set | Bullets | Lands | 
| --- | -------------- | -------------- | 
| 1 | `r unique(hamby224_set1$bulletA)` | `r unique(hamby224_set1$landA)` |
| 11 | `r unique(hamby224_set11$bulletA)` | `r unique(hamby224_set11$landA)` |

The dimensions of the datasets are:

| Set | Number of Rows | Number of Columns | 
| --- | -------------- | -------------- | 
| 1 | `r dim(hamby224_set1)[1]` | `r dim(hamby224_set1)[2]` |
| 11 | `r dim(hamby224_set11)[1]` | `r dim(hamby224_set11)[2]` |

The number of rows in each of these datasets indicates that there are some lands missing since a complete set with 3 bullets and 6 lands would contains 3 bullets x 3 bullets x 6 lands x 6 lands  = `r 6 * 6 * 3 * 3` comparisons. The figures below show that set 1 is missing land 4 from bullet Q, and set 11 is missing land2 from bullet 1 and land 4 from bullet I.

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, hjust = 1)),
  align = "hv"
)
```
Preparing the data from sets 1 and 11 for analysis:

```{r}
hamby224_set1_cleaned <-
  hamby224_set1 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features,-rfscore) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set1"),
    bullet1 = recode(factor(bullet1), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    bullet2 = recode(factor(bullet2), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource)

hamby224_set11_cleaned <- 
  hamby224_set11 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features,-rfscore) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set11"),
    bullet1 = recode(factor(bullet1), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    bullet2 = recode(factor(bullet2), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    land1 = str_remove(land1, "and "),
    land2 = str_remove(land2, "and ")
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource)
```

Plots show that the number of observations is still the same after cleaning:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Join the two cleaned Hamby 224 sets into one testing set, remove duplicate comparisons, create a case variable, and create land ids based on the study, set, bullet, and land (remove those variables afterwards):

```{r}
hamby224_test_extra_vars <-
  # Join the two data sets
  bind_rows(hamby224_set1_cleaned, hamby224_set11_cleaned) %>%
  # Remove duplicate comparisons
  filter(!(bullet1 == "BUnk" & bullet2 == "B1"),
         !(bullet1 == "BUnk" & bullet2 == "B2"),
         !(bullet1 == "B2" & bullet2 == "B1")) %>%
  # Create a case and land_id variables
  mutate(
    case = factor(1:length(study)),
    land_id1 = paste(study, set, bullet1, land1, sep = "-"),
    land_id2 = paste(study, set, bullet2, land2, sep = "-")
  ) # Leave the extract variables for plotting

# Select and order variables to keep for analysis
hamby224_test <- hamby224_test_extra_vars %>%
  select(case, land_id1, land_id2, all_of(rf_features), samesource)
```

Remaining comparisons included in the test data after the duplicate comparisons are removed:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_test_extra_vars %>%
    filter(set == "Set1") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_test_extra_vars %>%
    filter(set == "Set11") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Save the test data as a csv file:

```{r eval = FALSE}
write.csv(
  x = hamby224_test, 
  file = "../data/bullet-test.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/raw/h224-set1-features.rds.zip", 
  files = "../data/raw/h224-set1-features.rds"
)
zip(
  zipfile = "../data/raw/h224-set11-features.rds.zip", 
  files = "../data/raw/h224-set11-features.rds"
)
zip(
  zipfile = "../data/bullet-test.csv.zip", 
  files = "../data/bullet-test.csv"
)
```

# Example Matching Signatures

The following code trims the data provided by Heike of two matching bullet-land signatures

Import the data:

```{r}
signatures <- readRDS("../data/raw/signatures.rds")
```

Structure of the data:

```{r}
signatures %>% str()
```

Create a land variable based on the source variable, select only variables necessary for the manuscript, and rename sig as y:

```{r}
signatures_trimmed <-
  signatures %>%
  mutate(land = c("Signature 1", "Signature 2")[as.factor(source)]) %>%
  select(land, x, sig) %>%
  rename(y = sig)
```

Export the trimmed data as a csv file:

```{r eval = FALSE}
write.csv(
  x = signatures_trimmed,
  file = "../data/example-signatures.csv",
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/raw/signatures.rds.zip", 
  files = "../data/raw/signatures.rds"
)
zip(
  zipfile = "../data/example-signatures.csv.zip", 
  files = "../data/example-signatures.csv"
)
```