### 2020-09-25 LEAscans/bullete.Rmd
scans in md

---
output: rmarkdown::github_document
always_allow_html: yes
---

## The bullet that doesn't fit

Bullet E  of Hamby set 173 is given by NIST () as being fired from Barrel 1,   but visual inspection does not show a match to any other bullets fired from barrel 1 (known or questioned from either set).

Land 1

![Hamby173-BrE-B1-L1.png](Bullet E/Hamby173-BrE-B1-L1.png)

Land 2

![Hamby173-BrE-B1-L2.png](Bullet E/Hamby173-BrE-B1-L2.png)

Land 3

![Hamby173-BrE-B1-L3.png](Bullet E/Hamby173-BrE-B1-L3.png)

Land 4

![Hamby173-BrE-B1-L4.png](Bullet E/Hamby173-BrE-B1-L4.png)

Land 5

![Hamby173-BrE-B1-L5.png](Bullet E/Hamby173-BrE-B1-L5.png)

Land 6

![Hamby173-BrE-B1-L6.png](Bullet E/Hamby173-BrE-B1-L6.png)