### 2020-09-28 code/02-data-preparation.Rmd
some more work on updating training data for paper (but still working)

---
title: "Bullet Data Preparation for LIME Diagnostics Paper"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, fig.align = "center")
```

This document contains code for preparing the raw bullet data sets (training, testing, and example matching signatures) to be used the paper "Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations".

**NOTE** The code chunks where the prepared data versions are shared have `eval = FALSE`. These code chunks must be run manually. This is done to help avoid any accidental changes of the data.

Load R packages:

```{r}
library(cowplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(randomForest)
library(readr)
library(stringr)
library(tidyr)
library(zip)
```

Obtain features used when fitting the rtrees random forest:

```{r}
rtrees_features <- rownames(bulletxtrctr::rtrees$importance)
```

# Bullet Training Data

Load the training data provided by Heike (from CSAFE data base; see this [R script](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R) for the code used to access the data):

```{r}
if (!file.exists("../data/raw/CCFs_withlands.csv")) {
  unzip("../data/raw/CCFs_withlands.csv.zip", exdir = "../data/raw")
}
bullet_train_raw = read.csv("../data/raw/CCFs_withlands.csv")
```

The raw data contains `r dim(bullet_train_raw)[1]` observations (comparisons of signatures) and `r dim(bullet_train_raw)[2]` variables. The table below lists the levels of study, barrel, bullet, and land contained within the raw data. One of the study labels is `Hamby44`. CSAFE has determined that these bullets are actually from study `Hamby173`. Additionally, note that the barrel labels contains letters. These barrels are "unknown" barrels, and the new labeling convention used by CSAFE is to use the label for barrel of `BrUnk` and label the bullet using the letter (such as `BA`).

| Descriptor | Levels | 
| --- | -------------- |
| Study | `r unique(c(bullet_train_raw$study.x, bullet_train_raw$study.y))` |
| Barrel | `r unique(c(bullet_train_raw$barrel.x, bullet_train_raw$barrel.y))` |
| Bullet | `r unique(c(bullet_train_raw$bullet.x, bullet_train_raw$bullet.y))` |
| Land | `r unique(c(bullet_train_raw$land.x, bullet_train_raw$land.y))` |

The code below cleans and prepares the data for use in the paper. The steps taken are as follows:

1. Rename the label variables using numbers instead of letters (e.g. `study.x` and `study.y` become `study1` and `study2`, respectively). 
2. Select only the variables of interest for the study: signature ID labels (`study1`, `barrel1`, `bullet1`, `land1`, `study2`, `barrel2`, `bullet2`, and `land2`), features used to train the `rtrees` random forest model (`ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks`), and the variable identifying whether the two signatures are fired from the same gun (`same_source`).
3. Convert all categorical variables to characters.
4. Change the study labels to the correct studies and the barrel and bullet variable labels to the new CSAFE labeling conventions. 
5. Add letters to the barrel, bullet, and land labels to conform to the CSAFE labeling conventions.
6. Create land ID variables that combine study, barrel, bullet, and land labels. 
7. Add a case variable (just numbering comparisons 1 through `r dim(bullet_train_raw)[1]` based on current order in the data). 
8. Reorder the variables.

```{r}
# Identify the letters used to label barrels 
all_barrel_labels <- unique(c(bullet_train_raw$barrel.x, bullet_train_raw$barrel.y))
letters <- all_barrel_labels[!(all_barrel_labels %in% 1:10)]

bullet_train <- 
  bullet_train_raw %>%
  # 1. Rename variables
  rename(
    "study1" = "study.x", "barrel1" = "barrel.x", "bullet1" = "bullet.x", "land1" = "land.x",
    "study2" = "study.y", "barrel2" = "barrel.y", "bullet2" = "bullet.y", "land2" = "land.y",
    "same_source"= "match"
  ) %>%
  # 2. Select variables of interest
  select(
    study1, barrel1, bullet1, land1,
    study2, barrel2, bullet2, land2,
    all_of(rtrees_features), same_source
  ) %>%
  # 3. Convert categorical variables to characters
  mutate(across(c(barrel1, bullet1, land1, barrel2, bullet2, land2), as.character)) %>%
  # 4. Adjust study, barrel, and bullet labels
  mutate(
    study1 = ifelse(study1 == "Hamby44", "Hamby173", study1),
    study2 = ifelse(study2 == "Hamby44", "Hamby173", study2),
    bullet1 = ifelse(barrel1 %in% letters, barrel1, bullet1),
    bullet2 = ifelse(barrel2 %in% letters, barrel2, bullet2),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  # 5. Add abbreviations infront of barrel, bullet, and land labels
  mutate(
    barrel1 = paste0("Br", barrel1),
    barrel2 = paste0("Br", barrel2),
    bullet1 = paste0("B", bullet1),
    bullet2 = paste0("B", bullet2),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # 6. Create land ID variables
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  # 7. Add a case variable
  mutate(case = as.character(1:n())) %>%
  # 8. Reorder the variables
  select(land_id1, land_id2, 
         study1, barrel1, bullet1, study2, land1, barrel2, bullet2, land2,
         all_of(rtrees_features), same_source)
```

Check to make sure the number of rows in the data is still the same:

```{r}
dim(bullet_train)
```

The updated signature identification labels are included below:

| Descriptor | Labels | 
| --- | -------------- |
| Study | `r unique(c(bullet_train$study1, bullet_train$study2))` |
| Barrel | `r unique(c(bullet_train$barrel1, bullet_train$barrel2))` |
| Bullet | `r unique(c(bullet_train$bullet1, bullet_train$bullet2))` |
| Land | `r unique(c(bullet_train$land1, bullet_train$land2))` |

Determine the number of total study-barrel-bullet-lands in the data:

```{r}
length(unique(c(bullet_train$land_id1, bullet_train$land_id2)))
```

Save the prepared data as a csv file:

```{r eval = FALSE}
write.csv(
  x = bullet_train, 
  file = "../data/bullet-train.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/bullet-train.csv.zip", 
  files = "../data/bullet-train.csv"
)
```

# Bullet Testing Data

Load in the raw Hamby 224 datasets (sets 1 and 11; provided by Heike and CSAFE):

```{r}
if (!file.exists("../data/raw/h224-set1-features.rds")) {
  unzip("../data/raw/h224-set1-features.rds.zip", exdir = "../data/raw")
}
if (!file.exists("../data/raw/h224-set11-features.rds")) {
  unzip("../data/raw/h224-set11-features.rds.zip", exdir = "../data/raw")
}
hamby224_set1 <- readRDS("../data/raw/h224-set1-features.rds")
hamby224_set11 <- readRDS("../data/raw/h224-set11-features.rds")
```

The bullet and land names from each of the sets are included below. This indicates that each set contains `r length(unique(c(hamby224_set1$bulletA, hamby224_set1$bulletB)))` bullets with `r length(unique(c(hamby224_set1$landA, hamby224_set1$landB)))` lands each.

| Set | Bullets | Lands | 
| --- | -------------- | -------------- | 
| 1 | `r unique(c(hamby224_set1$bulletA, hamby224_set1$bulletB))` | `r unique(c(hamby224_set1$landA, hamby224_set1$landB))` |
| 11 | `r unique(c(hamby224_set11$bulletA, hamby224_set11$bulletB))` | `r unique(c(hamby224_set11$landA, hamby224_set11$landB))` |

The dimensions of the datasets are:

| Set | Number of Rows | Number of Columns | 
| --- | -------------- | -------------- | 
| 1 | `r dim(hamby224_set1)[1]` | `r dim(hamby224_set1)[2]` |
| 11 | `r dim(hamby224_set11)[1]` | `r dim(hamby224_set11)[2]` |

The number of rows in each of these datasets indicates that there are some lands missing since a complete set with 3 bullets and 6 lands would contains 3 bullets x 3 bullets x 6 lands x 6 lands  = `r 6 * 6 * 3 * 3` comparisons. The figures below show that set 1 is missing land 4 from bullet Q, and set 11 is missing land2 from bullet 1 and land 4 from bullet I.

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11 %>%
    count(bulletA, bulletB, landA, landB) %>%
    ggplot(aes(
      x = landA, y = landB, fill = n
    )) +
    geom_tile() +
    facet_grid(bulletB ~ bulletA) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1, 
          axis.text.x = element_text(angle = 45, hjust = 1)),
  align = "hv"
)
```

Preparing the data from sets 1 and 11 for analysis:

```{r}
hamby224_set1_cleaned <-
  hamby224_set1 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set1"),
    bullet1 = recode(factor(bullet1), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    bullet2 = recode(factor(bullet2), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rtrees_features), samesource)

hamby224_set11_cleaned <- 
  hamby224_set11 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set11"),
    bullet1 = recode(factor(bullet1), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    bullet2 = recode(factor(bullet2), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    land1 = str_remove(land1, "and "),
    land2 = str_remove(land2, "and ")
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rtrees_features), samesource)
```

Plots show that the number of observations is still the same after cleaning:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  hamby224_set1_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  hamby224_set11_cleaned %>%
    count(bullet1, bullet2, land1, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Join the two cleaned Hamby 224 sets into one testing set, remove duplicate comparisons, create a case variable, and create land ids based on the study, set, bullet, and land (remove those variables afterwards):

```{r}
bullet_test_extra_vars <-
  # Join the two data sets
  bind_rows(hamby224_set1_cleaned, hamby224_set11_cleaned) %>%
  # Remove duplicate comparisons
  filter(!(bullet1 == "BUnk" & bullet2 == "B1"),
         !(bullet1 == "BUnk" & bullet2 == "B2"),
         !(bullet1 == "B2" & bullet2 == "B1")) %>%
  # Create a case and land_id variables
  mutate(
    case = factor(1:length(study)),
    land_id1 = paste(study, set, bullet1, land1, sep = "-"),
    land_id2 = paste(study, set, bullet2, land2, sep = "-")
  ) # Leave the extra variables for plotting

# Select and order variables to keep for analysis
bullet_test <- bullet_test_extra_vars %>%
  select(case, land_id1, land_id2, all_of(rtrees_features), samesource)
```

Remaining comparisons included in the test data after the duplicate comparisons are removed:

```{r fig.width = 10, fig.height = 4.5}
plot_grid(
  bullet_test_extra_vars %>%
    filter(set == "Set1") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  bullet_test_extra_vars %>%
    filter(set == "Set11") %>%
    count(bullet1, land1, bullet2, land2) %>%
    ggplot(aes(
      x = land1, y = land2, fill = n
    )) +
    geom_tile() +
    facet_grid(bullet2 ~ bullet1) +
    scale_fill_distiller(palette = "GnBu", direction = 1) +
    theme_minimal() +
    theme(aspect.ratio = 1),
  align = "hv"
)
```

Save the test data as a csv file:

```{r eval = FALSE}
write.csv(
  x = bullet_test, 
  file = "../data/bullet-test.csv", 
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/bullet-test.csv.zip", 
  files = "../data/bullet-test.csv"
)
```

# Example Matching Signatures

The following code trims the data provided by Heike of two matching bullet-land signatures.

Import the data (provided by Heike and CSAFE):

```{r}
if (!file.exists("../data/raw/signatures.rds")) unzip("../data/raw/signatures.rds.zip", exdir = "../data/raw")
signatures <- readRDS("../data/raw/signatures.rds")
```

Structure of the data:

```{r}
signatures %>% str()
```

Create a land variable based on the source variable, select only variables necessary for the manuscript, and rename sig as y:

```{r}
signatures_trimmed <-
  signatures %>%
  mutate(land = c("Signature 1", "Signature 2")[as.factor(source)]) %>%
  select(land, x, sig) %>%
  rename(y = sig)
```

Export the trimmed data as a csv file:

```{r eval = FALSE}
write.csv(
  x = signatures_trimmed,
  file = "../data/example-signatures.csv",
  row.names = FALSE
)
```

Also, save the original and prepared data as a zip files (for uploading to the GitHub repository):

```{r eval = FALSE}
zip(
  zipfile = "../data/example-signatures.csv.zip", 
  files = "../data/example-signatures.csv"
)
```