### 2021-01-06 code/03-submission-steps.Rmd

---
title: "Sumbission Steps"
author: "Katherine Goode"
date: "<br>`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, eval = FALSE)
```

This document contains the steps to follow to make sure all guidelines are met for submission to the ASA Data Science Journal.

**NOTE** The knitr option of `eval = FALSE` has been set for this document, so all code must be run manually from within the Rmd file. This is done to be safe about not deleting or replacing files.

0. Run the code if desired to update the saved state of the project R package versions. Use with caution!

    ```{r}
    # Save the r package versions
    renv::snapshot()
    ```

1. Delete all data files created from the raw versions, files generated by `paper.Rnw`, and current existing submission folder using the code below.

    ```{r}
    # Specify date of submission folder
    submission_date = "2021-01-06"

    # Create a submission folder file path:
    submission_folder <- paste0("../submissions/", submission_date)

    # Individual files to delete
    file.remove("../data/bullet-explain.rds")
    file.remove("../data/bullet-explain-perms-clean.rds")
    file.remove("../data/bullet-lime.rds")
    file.remove("../data/bullet-rf.rds")
    file.remove("../data/bullet-test.csv")
    file.remove("../data/bullet-test.csv.zip")
    file.remove("../data/bullet-train.csv")
    file.remove("../data/bullet-train.csv.zip")
    file.remove("../data/example-signatures.csv")
    file.remove("../data/example-signatures.csv.zip")
    file.remove("../data/sine-data-test.rds")
    file.remove("../data/sine-data-train.rds")
    file.remove("../data/sine-lime-explain.rds")
    file.remove("../data/raw/h224-set1-features.rds")
    file.remove("../data/raw/h224-set1-features.rds.zip")
    file.remove("../data/raw/h224-set11-features.rds")
    file.remove("../data/raw/h224-set11-features.rds.zip")
    file.remove("../data/raw/CCFs_withlands.csv")
    file.remove("../data/raw/CCFs_withlands.csv.zip")
    file.remove("../data/raw/signatures.rds")
    file.remove("../data/raw/signatures.rds.zip")
    file.remove("../figure-static/figure-06-1.eps")
    file.remove("../figure-static/figure-06-1.png.zip")
    file.remove("../figure-static/figure-08-1.eps")
    file.remove("../figure-static/figure-08-1.png")

    # Folders to delete
    unlink("../figure", recursive = TRUE)
    unlink(submission_folder, recursive = TRUE)
    unlink("../support-info/figure", recursive = TRUE)
    ```

2. Manually run the code in `code/01-raw-file-compression.Rmd` to add compressed versions of the raw data files and to compress the static figure for the manuscript. Also, knit the file to make sure the md version is up to date.

3. Manually run the code in `code/02-data-preparation.Rmd` to create the versions of the data to be used in the manuscript and submitted with the manuscript. Also, knit the file to make sure the md version is up to date.

4. Knit the file `paper.Rnw`.

5. Knit the supporting information document `support-info/supporting-information.Rnw`.

6. Run the following code to move all files necessary to the submission folder.

    ```{r}
    # Load package
    library(zip)

    # If the submission folder does not exist, create one
    if (!dir.exists(submission_folder)) dir.create(submission_folder)

    # DATA AND CODE

    # Create a temporary directory for storing code and data files for submission
    code_data_folder = "code-and-data"
    dir.create(code_data_folder)

    # Extract the R code from the files paper.Rnw to be used for submission
    knitr::purl("../paper.Rnw", output = paste0(code_data_folder, "/code.R"))

    # Create a vector with code and data files to be included for submission 
    data_files <-
      c(
        "../data/README.txt",
        "../data/bullet-test.csv",
        "../data/bullet-train.csv",
        "../data/example-signatures.csv"
      )

    # Copy the submission files to the temporary folder
    file.copy(data_files, code_data_folder)

    # Compress the folder with data and code for submission and store it in the 
    # submission folder
    zip(paste0(submission_folder, "/code-and-data.zip"), code_data_folder)

    # Remove the temporary folder with code and data for submission
    unlink(code_data_folder, recursive = TRUE)

    # FIGURES

    # Identify the folders with files for submission
    folder_eps_figures <- "../figure"
    folder_static_figures <- "../figure-static"

    # Find the EPS figure files for submission
    list_of_generated_figures <-
      list.files(folder_eps_figures, ".eps$", full.names = T)
    list_of_static_figures <-
      list.files(folder_static_figures, ".eps$", full.names = T)

    # Copy files to the submission folder
    file.copy(list_of_generated_figures, submission_folder)
    file.copy(list_of_static_figures, submission_folder)

    # LATEX FILES

    # Copy files to the submission folder
    file.copy("../paper.tex", submission_folder)
    file.copy("../paper.pdf", submission_folder)
    file.copy("../references.bib", submission_folder)
    file.copy("../WileyNJD-AMS.bst", submission_folder)
    file.copy("../WileyNJD-v2.cls", submission_folder)

    # SUPPORTING INFORMATION

    # Copy the supporting information pdf to the submission folder
    file.copy("../support-info/supporting-information.pdf", submission_folder)
    ```

7. Manually delete `figure/` and `figure-static/` from the figure file paths in `paper.tex`. For example, change `\includegraphics[width=6.5in]{figure/figure-01-1}` to `\includegraphics[width=6.5in]{figure-01-1}`.

8. Compile the file `paper.tex` to make sure it compiles correctly (especially references). (If using texShop, make sure to compile using LaTex, then bibTex, then LaTeX twice more.)

9. Besides for `paper.pdf`, delete all other files generated when `paper.tex` was compiled. This includes other paper files and pdf versions of the figures.

    ```{r}
    # Remove the paper files that may have been generated during compilation
    file.remove("../submission/paper.aux")
    file.remove("../submission/paper.bbl")
    file.remove("../submission/paper.blg")
    file.remove("../submission/paper.log")
    file.remove("../submission/paper.synctex.gz")
    file.remove("../submission/paper.pag")

    # Remove the pdf versions of the figures generated during compilation
    file.remove(list.files(
      path = "../submission/",
      pattern = "converted-to",
      full.names = TRUE
    ))
    ```

10. Upload all remaining files in the submission folder to the ASA Data Science Journal submission page which should be:

  - `code-and-data.zip`
  - `figure-01-1.eps` through `figure-B1-1.eps`
  - `paper.pdf`
  - `paper.tex`
  - `references.bib`
  - `supporting-information.pdf`
  - `WileyNJD-AMS.bst`
  - `WileyNJD-v2.cls`

10. Knit this document to make sure the md version is up to date.